Comparing microbial communities with OptiFit: a fast method for fitting sequences to existing OTUs

### 2021-03-11 docs/slides/2021-03-11_bistro.Rmd
Add lots of logos

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#23395b",
  header_font_google = google_font("Helvetica"),
  text_font_google = google_font("Helvetica")
)
```

@@

@@

.center[
<img src="https://raw.githubusercontent.com/kelly-sovacool/bioinf603-fall2019/master/docs/figures/human_microbiome.gif" height="500"/>
]
---

@@

|   | Metagenomics | Amplicon Sequencing |
|------|--------------|----------|
| **Technique** | Whole-metagenome shotgun sequencing (every gene of every microbe) | Sequence just 250bp of the 16S rRNA gene |
| **Taxonomic resolution** | Species or strain | Genus |
| **Organisms** | Bacteria, archaea, viruses, fungi | Bacteria & archaea; or fungi separately | 
| **Purpose** | Identify which microbial genes are present; assemble microbial genomes | Identify which microbes are present and their abundances |
| **Cost** | 💲💲💲 | 💲 |

@@

- Cluster similar sequences together into OTUs.
- OTUs clustered at 3% distance (97% sequence similarity) generally can be labelled at the genus level.

@@

@@

.center[
<img src="../../figures/opticlust-supp-ex.png" width="700"/>
]
<img src="https://raw.githubusercontent.com/mothur/logo/master/mothur_RGB.png" width="200"/>

???
if equally good, pick one at random

@@

## Reproducible Pipelines with Snakemake

@@

@@

.pull-right[
<img src="../../figures/dag_fit_db.png" height="600" />
]

>>>>

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  #base_color = "#23395b",
  base_color = "#000000",
  header_font_google = google_font("Helvetica"),
  text_font_google = google_font("Helvetica")
)
```

@@

background-image: url("../../figures/SchlossLab.png") 
background-size: contain
background-color: "#000000"
class: inverse

@@

.center[
<img src="https://raw.githubusercontent.com/kelly-sovacool/bioinf603-fall2019/master/docs/figures/human_microbiome.gif" height="500"/>
]

@@

|   | Metagenomics | Amplicon Sequencing |
|------|--------------|----------|
| **Technique** | Whole-metagenome shotgun sequencing: every gene of every microbe | Sequence just 250bp of the 16S rRNA gene |
| **Cost** | 💲💲💲 | 💲 |
| **Taxonomic resolution** | Species or strain | Genus |
| **Organisms** | Bacteria, archaea, viruses, fungi | Bacteria & archaea; or fungi separately | 
| **Purpose** | Identify which microbial genes are present; assemble microbial genomes | Identify which microbes are present and their abundances |

???
metagenomics is prohibitively expensive for very large datasets

@@

- Cluster similar sequences together into OTUs as a proxy for genus/species.
- OTUs clustered at 3% distance (97% sequence similarity) generally can be labelled at the genus level.

@@

Courtney's resolution figure?

@@

.center[
<img src="../../figures/opticlust-supp-ex.png" width="700"/>
]
.footnote[
<img src="https://raw.githubusercontent.com/mothur/logo/master/mothur_RGB.png" width="200"/>
]
???
if equally good, pick one at random

@@

## Managing Bioinformatics Pipelines

@@

.footnote[
[snakemake.readthedocs.io](snakemake.readthedocs.io)
]

???
I ran about 20,000 jobs on the GreatLakes HPC

@@

## Pipelines with <img src="https://raw.githubusercontent.com/snakemake/snakemake/f55801bfabfb916f77fa4ec8f9eb0667e5c8a92e/images/biglogo.png" height="35"/>

.center[
<img src="https://raw.githubusercontent.com/snakemake/snakemake/f55801bfabfb916f77fa4ec8f9eb0667e5c8a92e/docs/project_info/img/idea.png" width="900" />
]

.footnote[
[snakemake.readthedocs.io](snakemake.readthedocs.io)
]

## Tools for Reproducible Bioinformatics

| package mgmt | scripts | reports | workflow | version control |
|---|---|---|---|---|
|   <img src="https://docs.conda.io/en/latest/_images/conda_logo.svg" width="175"/> | <img src="https://longhowlam.files.wordpress.com/2017/04/rp.png" height="175"/> | <img src="https://ulyngs.github.io/rmarkdown-workshop-2019/slides/figures/rmarkdown.png" height="165"/> | <img src="https://avatars.githubusercontent.com/u/33450111?s=400&v=4" height="175"> |<img src="https://miro.medium.com/max/2625/1*8fPMdk2Cd5iJQ7dI7jXCbA.jpeg" height="175"/> |

.center[
### ...and headache prevention!
]
